---
title: "applyNA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{applyNA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(epiuf)

```


# Apply NA function

The aim of this function is to take a variable and set specified values to NA. Uses gsub to acheive this. Values to search can be specified in list, or in string format. Usual nomenclature applies to specify beginning, end, ect. A default list is available that is used if no values are specified or that can be join to the specified value list.


```{r dataload}

# create example dataset for testing
df <- data.frame(CharacterVar = c("unknown", "do not know", "other", "dnk", "na", "nd", "nsp", "ne zna", "testdnk", "dnktest", "test", "not test", "3", "4"), 
                 NumberVar = c(1, 2, 3, 4, 1, 2, 3, 4,1, 2, 3, 4, 1, 2)
)


```


# Testing the first draft of the function

```{r testfunction}

df$test <- applyNA(df, CharacterVar)  # only default list
# Should make 7 conversions

df$test <- applyNA(df, CharacterVar, "^test$") # only specified
# Should make 1 conversions  

df$test <- applyNA(df, CharacterVar, "^test$", join=TRUE) # both default and specified
# Should make 8 conversions  


df$test <- applyNA(df, CharacterVar, c("^test$","^other*"))       # list of search values
# Should make 2 conversions  


df$test <- applyNA(df, NumberVar, 3)        # number value
# Should make 3 conversions

df$test <- applyNA(df, NumberVar, c(3,1))         # list of numbers
# Should make 7 convers


```

