---
title: "collapseVar"
author: "JHD"
date: "24/03/2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{collapseVar}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(epiuf)


```

collapseVar take number list variables in character format with any separator (eg. "1|2|3", "1/2/3", "1,2,3"), and retains only one input / row based off hirearchy specified and return numeric variable.
If order input is not complete, a warning is given.


```{r testdata}

# create example dataset for testing
df <- data.frame(checkbox_bar = c("0|2", "1|2|3", "3|1|2", "0", "2", "3", "0|1", "2|3", "")
                 ,checkbox_comma = c("0,2", "1,2,3", "3,1,2", "0", "2", "3", "0,1", "2,3", "")
                 ,checkbox_slash = c("0/2", "1/2/3", "3/1/2", "0", "2", "3", "0/1", "2/3", "")
)

```


# Test ability to handle different separators
```{r tests1}
df$test <- collapseVar(df, checkbox_bar, c(0,1,2,3))
df$test <- collapseVar(df, checkbox_comma, c(0,1,2,3))
df$test <- collapseVar(df, checkbox_slash, c(0,1,2,3))
``` 

Expected outcome for each:
0  1  1  0  2  3  0  2 NA


# Test what happens if extra numbers in input
```{r tests2}
df$test <- collapseVar(df, checkbox_bar, c(0,1,2,3,4))
``` 
Expected outcome:
0  1  1  0  2  3  0  2 NA

# Test what happens if insufficient numbers in input
```{r tests3}
df$test <- collapseVar(df, checkbox_bar, c(0,2))
``` 
Expected outcome:
Warning: input order list incomplete, missing 1 3
