---
title: "dataCheck"
author: "JHD"
date: '2022-06-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(epiuf)
```

checkDico takes a dataset, a condition and a caption. It extracts the rows which meet the conditition, and ouputs the caption, the count of rows, and (if less the threshold) the list of or a specified variable, usually id. An optional extra, if 2 datasets are input, repeated IDs meeting the criteria are highlighted by [ ] so that new IDs can be quickly identified. 


```{r testdata}
df1 <- data.frame(Id = c(1,2,3,4,5,6,7,8,9,10)
                  , A = c("a", "b", "c", "d", "e","a", "b", "c", "d", "e")
                  , B = c("", "dog", "cat", "rabbit", "mole", "dog", "horse", "cat", "", "")
                  )
df2 <- data.frame(Id = c(6,7,8,9,10, 11,12,13,14,15)
                  , A = c("a", "b", "c", "d", "e","f", "b", "c", "d", "e")
                  , B = c("dog", "horse", "cat", "", "mouse","mole", "antelope", "antelope", "hamster", "")
                  )

```

Currently, when want to only pass 1 data frame, must specify data_old=NULL
```{r tests1}
dataCheck(data=df1, data_old=NULL, cond = A=="b", "B's in df1 only", varname="Id")
# output is 7 and 2
dataCheck(data=df1, data_old=df2, cond = A=="b", "B's in df1 and df2", varname = "Id")
# output is 7 in both, 2 in df1 only
```
```{r tests2}
dataCheck(data=df1, data_old=NULL, cond = A=="b"&B=="mouse", "B's and mice in df1 only", varname="Id")
# output should be null
dataCheck(data=df1, data_old=df2, cond = A=="b"&B=="mouse", "B's and mice in df1 and df2", varname = "Id")
# output should be null

dataCheck(data=df1, data_old=df2, cond = A=="b"&B=="horse", "B's and horses in df1 and df2", varname = "Id")
# output should be 7 from both.
```
