---
title: "dataCheck"
author: "JHD"
date: '2022-06-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(epiuf)
```

The data check suite: core functions to extract each element of a data check, and wrapper function to generate outputs in each format we 

countIf() : the count of records matching a given condition
listIf() : the list of specified variable for records matching a given condition
printIf() : the count of records and list if under a specified threshold
dataCheck() takes a dataset, a condition and a caption. It extracts the rows which meet the conditition, and ouputs the caption, the count of rows, and (if less the threshold) the list of or a specified variable, usually id. An optional extra, if 2 datasets are input, repeated IDs meeting the criteria are highlighted by [ ] so that new IDs can be quickly identified. 
Attention: Must specify each parameter.


```{r testdata}
df1 <- data.frame(Id = c(1,2,3,4,5,6,7,8,9,10, NA_real_)
                  , A = c("a", "b", "b", "d", "b","a", "b", "b", "d", "e", "b")
                  , B = c("", "dog", "cat", "rabbit", "mole", "dog", "horse", "cat", "", NA_character_, "mouse")
                  )
df2 <- data.frame(Id = c(6,7,8,9,10, 11,12,13,14,15, NA_real_)
                  , A = c("a", "b", "b", "d", "b","f", "b", "b", "d", "e", "b")
                  , B = c("dog", "horse", "cat", "", "mouse","mole", "antelope", "antelope", "hamster", "", "mouse")
                  )

```

```{r countiftest}
countIf(data=df1, cond = A=="b")
countIf(df1, A=="b")

```

```{r listiftest}
listIf(data=df1, varname="Id", cond = A=="b")
listIf(data=df1, varname="Id", cond = A=="b", na.rm=TRUE)

listIf(df1, "Idcheck", A=="b")
listIf(df1, "Idcheck")
listIf(df1, "Id", collapse=TRUE, na.rm=TRUE)

```

```{r printIf2tests}
printIf2(data=df1, cond = A=="b", text = "B's in df1 only", varname="Id")
# output is 7 and 2
printIf2(data=df1, data_old=df2, cond = A=="b",text = "B's in df1 and df2", varname = "Id")
# output is 7 in both, 2 in df1 only
```
```{r tests2}
dataCheck(data=df1, cond = A=="b"&B=="mouse", text = "B's and mice in df1 only", varname="Id")
# output should be null
dataCheck(data=df1, data_old=df2, cond = A=="b"&B=="mouse", text = "B's and mice in df1 and df2", varname = "Id")
# output should be null

dataCheck(data=df1, data_old=df2, cond = A=="b"&B=="horse", text = "B's and horses in df1 and df2", varname = "Id")
# output should be 7 from both.
```
