---
title: "dictionarytools"
author: "epi-gde"
date: "1/31/2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dictionarytools}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, results = "asis")
library(epiuf)

```

## Update source to dictionary generic structure

This is a first rename function which use a data dictionary.

```{r dico3}

source <- data.frame(source1=c(1,2,3),source2=c(2,3,4),source3=c(4,5,6),source6=c(6,7,8))

dic <- data.frame(generic_name=c("gen1","gen2","gen3","gen4","gen5",NA),
                  source_name=c("source1","source2","source3",NA,NA,"source6"),
                  type=c("numeric","numeric","character","numeric","character",NA))


new <- applyDictionary(dictionary=dic,data=source)


```

## Test for error (duplicated variable name in dico)

This is a test of the data dictionary. duplicated names are not allowed in the dictionnary generic variable list

```{r dicodup }

source <- data.frame(source1=c(1,2,3),source2=c(2,3,4),source3=c(4,5,6),source6=c(6,7,8))

dic <- data.frame(generic_name=c("gen1","gen1","gen3","gen4","gen5",NA),
                  source_name=c("source1","source2","source3",NA,NA,"source6"),
                  type=c("numeric","numeric","numeric","character","character",NA))

new <- applyDictionary(dictionary=dic,data=source)


```

## Verify Nothing to do

This is a first rename function which use a data dictionary.

```{r nothingtodo }

source <- data.frame(source1=c(1,2,3),source2=c(2,3,4),source3=c(4,5,6))

dic <- data.frame(generic_name=c("source1","source2","source3"),
                  source_name=c("source1","source2","source3"),
                  type=c("numeric","numeric","numeric"))

new <- applyDictionary(dictionary=dic,data=source)


```

## Keep Extra variables in updated data

```{r dico4}

source <- data.frame(source1=c(1,2,3),source2=c(2,3,4),source3=c(4,5,6),source4=c(5,6,7), source6=c(6,7,8))

dic <- data.frame(generic_name=c("gen1","gen2","gen3","gen4","gen5",NA),
                  source_name=c("source1","source2","source3",NA,NA,"source6"),
                  type=c("numeric","numeric","numeric","character","character",NA))


new <- applyDictionary(dic,source,keepextra = TRUE)


```

## Complex version which mix everything

```{r dico5}

source <- data.frame(source1=c(1,2,3),source2=c(2,3,4),source4=c(4,5,6),source6=c(6,7,8))

dic <- data.frame(generic_name=c("gen1","gen2","gen3","gen4","gen5",NA,NA),
                  source_name=c("source1","source2","source3","",NA,"source6",NA),
                  type=c("numeric","numeric","numeric","character"," ","character",NA))


new <- applyDictionary(dic,source)

```

------------------------------------------------------------------------

## Utility function for dico

## Retrieve values from dico by name

```{r dico6, ECHO = FALSE}

dic <- data.frame(generic_name=c("gen1","gen2","gen3","gen4","gen5",NA,NA),
                  source_name=c("source1","source2","source3","",NA,"source6",NA),
                  type=c("numeric","numeric","numeric","character"," ","character",NA),
                  unknowns=c("","8","UNK","","","",""))

setDictionary(dictionary = dic)

val <- getDictionaryValue("gen1","type")
print(val)

val <- getDictionaryValue("gen3","unknowns")
print(val)

```

## Retrieve the dico dataset

```{r dico7, ECHO = FALSE}

dic <- data.frame(generic_name=c("gen1","gen2","gen3","gen4","gen5",NA,NA),
                  source_name=c("source1","source2","source3","",NA,"source6",NA),
                  type=c("numeric","numeric","numeric","character"," ","character",NA)
                  )

setDictionary(dictionary = dic)

retrieved <- getDictionary()
print(retrieved)

```

## Manage dico

```{r dico8, ECHO = FALSE}

dic <- data.frame(generic_name=c("gen1","gen2","gen3","gen4","gen5",NA,NA),
                  source_name=c("source1","source2","source3","",NA,"source6",NA),
                  type=c("numeric","numeric","numeric","character"," ","character",NA),
                  unknowns=c("","8","UNK","","","","")
                  )

setDictionary(dictionary = dic)
FileName <-   "generic dictionary.xlsx" 
saveDictionary(filename = FileName)

if (file.exists(FileName)) {print("Dictionary successfully saved")}


```

## Manage dico open file

```{r dico9, ECHO = FALSE}

FileName <-   "generic dictionary.xlsx" 
openDictionary(filename = FileName)

getDictionary()

# clean dir 
file.remove(FileName)

```

## Manage dico

```{r dico10, ECHO = FALSE}

dic <- data.frame(generic_name=c("gen1","gen2","gen3","gen4","gen5",NA,NA),
                  source_name=c("source1","source2","source3","",NA,"source6",NA),
                  type=c("numeric","numeric","numeric","character"," ","character",NA),
                  unknowns=c("","8","UNK","","","","")
                  )

setDictionary(dictionary = dic)
OneLine <- getNewDictionaryLine()
print(OneLine)



```

```{r dicovalue, ECHO = FALSE}

dic <- data.frame(generic_name=c("gen1","gen2","gen3","gen4","gen5",NA,NA),
                  source_name=c("source1","source2","source3","",NA,"source6",NA),
                  type=c("numeric","numeric","numeric","character"," ","character",NA),
                  unknowns=c("","8","UNK","","","","")
                  )

setDictionary(dictionary = dic)
getDictionaryValue("gen3","unknowns")
getDictionaryValue("gen1","unknowns")
getDictionaryValue("gen9","unknowns")
tryCatch(getDictionaryValue("gen3","unknows"))

```
