---
title: "findAndReplace"
author: "epi-gde"
date: "2023-10-20"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{findAndReplace}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(epiuf)

```

## **filesFindReplace**

This function allows to find a character, string or word into a file, a
series of files or a directory (with optional file filter) and to
replace it by another character, string or word.

R script, text files and csv can be processed as well as xlsx files.

Some functions have been used to build that functionality

The **charCount** function will be used to count changes before applying
them

```{r charCount}
  # A single character 
  charCount("c","cancan")
  charCount("c","ccc" )
  # A string in a sentence 
  charCount("essai"," this is an essai" )
  # A string in a vector 
  charCount("ce",c("cacec","ceci" ))
  # A string in a dataframe 
  df <- data.frame(c("cacececi","ceci" ,"non","ace"))
  charCount("ce",df)
  # A word in a dataframe  (between \b which escaped is \\b)
  charCount("\\bce\\b",df)
  df <- data.frame(c("ca cece ci","ce ci" ,"non","a ce"))
  charCount("\\bce\\b",df)
  df <- data.frame(c("ca cece ci","ce ci" ,"non","a ce"),c("ca cececi","ce ci" ,"non","ace"))
  charCount("\\bce\\b",df)
  # Verify that nothing is 0 
  charCount("","essai")
  charCount("e","")
  # A special character .
  charCount("\\.","21.15.98")
  
  
```

```{r prepFiles, echo=FALSE}
# prepare files for tests, this method avoid to store file in the package
# created files will be delete at the end of the process
file1 <- "temp_text1.R"
file2 <- "temp_text2.R"
file3 <- "temp_text3.R"
lines <- c(
"# test script for findAndReplace",
"# word to be replaced : oldword",
"#  old_word",
"#  oldword --  oldword)")

writeLines(lines, file1)
writeLines(lines, file2)
writeLines(lines, file3)
```

**listFiles** is a simplified version of list.files which accept jokers
of global search \* for any character and ? for one character (fullpath
is automaticaly set with recursive flag)

```{r listFiles}
listFiles(pattern="temp_*.R")
listFiles("R","c*")
```

**externalFile** is used for test and vignettes. In the package, you can
add some data/files These are stored in a specific directory attached to
the package like :
"/Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library/epiuf/extdata/testrename.xlsx"
But while you are building the package, the file is into inst/extdata.
Retrieving that file could be tricky from vignettes or test depending
your knit directory. externalFile while always return the right path
either in build mode or library mode, and whathever is your knit
directory

externalFile("testrename.xlsx")
"/Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library/epiuf/extdata/testrename.xlsx"

externalFile("testrename.xlsx") "/Users/gedeonz/Dev/R
Sources/STRAP-epiuf/inst/extdata/testrename.xlsx"

**filesFindReplace** is a generic function to make batch replacement You
can call this function for one file, a list of files, a directory, a
pattern to filter files or any combination of previous option. In
addition you can replace a list of pattern by one single replacement The
listonly parameters allow to check how many changes will be done.

```{r replaceinFiles}
  # replace in 2 files (count only)
  filesFindReplace(c(file1,file2),pattern="oldword",replacement = "newword",listonly=TRUE)
  # replace 2 pattern at a time (count only) 
  filesFindReplace(file3,pattern=c("oldword","old_word"),replacement = "newword",listonly=TRUE)
  # replace by modifying the original file
  filesFindReplace(file1,pattern="oldword",replacement = "newword")
  cat(readLines(file1),sep="\n")
  # replace in a directory or use a file filter
  filesFindReplace("temp_*.R",pattern="oldword",replacement = "newword")
  
```

**filesFindReplace** work on text files (R, txt, csv ...) and xlsx
files.

```{r replaceinxlsxFiles}
  # we use a file stored in packages data 
  file4 <- externalFile("testrename.xlsx")
  file5 <- "temp.xlsx"
  # In ordre to not modify the original file, we copy it to a temporary file which will be deleted at the end
  file.copy(file4,file5)
  # Count the number of occurrence in file
  filesFindReplace(file5,"patient","subject",word=TRUE, ignore.case = TRUE, listonly=TRUE)
  # Replace in file
  filesFindReplace(file5,"patient","subject",word=TRUE, ignore.case = TRUE)
  # Number of occurrence should be 0 after changes 
  filesFindReplace(file5,"patient","subject",word=TRUE, ignore.case = TRUE, listonly=TRUE)
  # we can manage more than one file filter 
  filesFindReplace(c("temp_*.R","temp*.xlsx"),pattern="oldword",replacement = "newword", listonly = TRUE)

 
```

```{r cleaning, echo=FALSE }
if(file.exists(file1)) r <- file.remove(file1)
if(file.exists(file2)) r <- file.remove(file2)
if(file.exists(file3)) r <- file.remove(file3)
if(file.exists(file5)) r <- file.remove(file5)

```
