---
title: "findAndReplace"
author: "epi-gde"
date: "2023-10-20"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{findAndReplace}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(epiuf)

```

The **charCount** function will be used to count changes before applying them

```{r charCount}
  # A single character 
  charCount("c","cancan")
  charCount("c","ccc" )
  # A string in a sentence 
  charCount("essai"," this is an essai" )
  # A string in a vector 
  charCount("ce",c("cacec","ceci" ))
  # A string in a dataframe 
  df <- data.frame(c("cacececi","ceci" ,"non","ace"))
  charCount("ce",df)
  # A word in a dataframe  (between \b which escaped is \\b)
  charCount("\\bce\\b",df)
  df <- data.frame(c("ca cece ci","ce ci" ,"non","a ce"))
  charCount("\\bce\\b",df)
  df <- data.frame(c("ca cece ci","ce ci" ,"non","a ce"),c("ca cececi","ce ci" ,"non","ace"))
  charCount("\\bce\\b",df)
  # Verify that nothing is 0 
  charCount("","essai")
  charCount("e","")
  # A special character .
  charCount("\\.","21.15.98")
  
  
```

```{r prepFiles, echo=FALSE}
# prepare files for tests
file1 <- "temp_text1.R"
file2 <- "temp_text2.R"
file3 <- "temp_text3.R"
lines <- c(
"# test script for findAndReplace",
"# word to be replaced : oldword",
"#  old_word",
"#  oldword --  oldword)")

writeLines(lines, file1)
writeLines(lines, file2)
writeLines(lines, file3)
```

**listFiles** is a simplified version of list.files which accept jokers of global search
* for any character and ? for one character
(fullpath is automaticaly set with recursive flag)

```{r listFiles}
listFiles(pattern="temp_*.R")
listFiles("R","c*")
```

**filesFindReplace** is a generic function to make batch replacement You can call this function for one file, a list of files, a directory, a pattern to filter files or any combination of previous option. In addition you can replace a list of pattern by one single replacement
The listonly parameters allow to check how many changes will be done. 

```{r replaceinFiles}
  # replace in 2 files (count only)
  filesFindReplace(c(file1,file2),pattern="oldword",replacement = "newword",listonly=TRUE)
  # replace 2 pattern at a time (count only) 
  filesFindReplace(file3,pattern=c("oldword","old_word"),replacement = "newword",listonly=TRUE)
  # replace by modifying the original file
  filesFindReplace(file1,pattern="oldword",replacement = "newword")
  cat(readLines(file1),sep="\n")
  # replace in a directory or use a file filter
  filesFindReplace("temp_*.R",pattern="oldword",replacement = "newword")
  
``` 

**filesFindReplace** work on text files (R, txt, csv ...) and xlsx files.


```{r replaceinxlsxFiles}
  
  file4 <- externalFile("testrename.xlsx")
  file5 <- "temp.xlsx"
  file.copy(file4,file5)
  # Count the number of occurrence in file
  filesFindReplace(file5,"patient","subject",word=TRUE, ignore.case = TRUE, listonly=TRUE)
  # Replace in file
  filesFindReplace(file5,"patient","subject",word=TRUE, ignore.case = TRUE)
  # Number of occurrence should be 0 after changes 
  filesFindReplace(file5,"patient","subject",word=TRUE, ignore.case = TRUE, listonly=TRUE)
 
```

```{r cleaning, echo=FALSE }
if(file.exists(file1)) r <- file.remove(file1)
if(file.exists(file2)) r <- file.remove(file2)
if(file.exists(file3)) r <- file.remove(file3)
if(file.exists(file5)) r <- file.remove(file5)

```
