---
title: "verifySpelling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{verifySpelling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r , setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(epiuf)
library(kableExtra)

```

## Epifunction : verifySpelling 

This document show examples of epiFunction use
```{r, config, echo = FALSE, include=FALSE}
ext_file <- externalFile("testbrand.csv")
# file is created in UTF-8 
df <- read.csv(ext_file,fileEncoding="UTF-8")

```

We use a brand name list to compare incorrectly spelled values

```{r, verifySpelling , echo = TRUE}

BrandName <- c("Pfizer","Moderna","Comirnaty","AstraZeneca","Spikevax",
                     "Jansen","Comirnaty (Pfizer/BioNTech)","vaxzevria",
                     "Johnson & Johnson","Pfizer/BioNTech Comirnaty")

verifySpelling("Maderna", BrandName)

```

We can use verifySpelling with sapply to add a new column containing guessed correct value in the dataset 
```{r, addColumn , echo = TRUE}

df$correctBrand <- verifySpelling(df$covvaccbrand_firstdose,BrandName,ErrPerc=0.4)
head(df,10)
```
We can also use a named list of brand name to regroup value like a recode 
(But that recode accept similar values instead of equal values)

```{r, verify and recode , echo = TRUE}

BrandList <- c("Pfizer"="Pfizer",
               "Pfizer"="Pfizer/BioNTech Comirnaty",
               "Pfizer"="Comirnaty (Pfizer/BioNTech)",
               "Pfizer"="Comirnaty",
               "Moderna"="Moderna",
               "Moderna"="Spikevax",
               "AstraZeneca"="AstraZeneca",
               "AstraZeneca"="Vaxzevria",
               "Janssen"="Janssen",
               "Johnson & Johnson"="Johnson & Johnson"
               )


df$recodedBrand <- verifySpelling(df$covvaccbrand_firstdose,BrandList,ErrPerc=0.4)
head(df,10)


```

All values which are not matching are NA so we can check them 

```{r, checkError , echo = TRUE}

dfError <- df[is.na(df$correctBrand),]
knitr::kable(dfError,row.names = FALSE) 
freq(dfError$recodedBrand,missing=TRUE)

```



Result saved to :
```{r, saveresult , eval=FALSE, echo = FALSE}
write.csv(dfError,file="vignettes/Brand Error.csv")

```






