---
title: "xlsxUtils"
author: "epi-gde"
date: "1/24/2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{xlsxUtils}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(epiuf)

```

## xlsxUtility functions

Using xlsx to write simple values to cells may be repetitive and trivial. This set of functions should simplify a lot the process

The first avalaible function is **openXlsx** This function load a workbook in memory The workbook is a java object and having it in environment does help so much... Then this workbook is stored into the functions them self

It is also possible to assign the workbook to a local variable if needed (to work with more than one workbook at a time)

```{r data}
xls_file <- externalFile("excelfile.xlsx")
openXlsx(xls_file)

```

## the openSheet() function

This function get the content of one sheet from the previously loaded workbook

```{r usage }

cells <- openSheet("T2")
```

Optionnaly you can load cells from a local workbook

```{r usage2 }
xls_file <- externalFile("excelfile.xlsx")
wb <- openXlsx(xls_file)
cells <-  openSheet("T2",wb)

```

And finally the function fillCells can be used to input values into a row of cells

```{r full exemple }
num <-  67
denum <-  90 
cells <- openSheet("T2")
fillCells(cells, 2 , 2 , "Cases","Perc")
fillCells(cells, 3 , 2 , num, num/denum )
cells <- openSheet("T3")
fillCells(cells, 4 , 2 , "total",100)


```

```{r matrice}
mat <- data.frame(Id = 1:3 , Vaccs = c("1", "3", "6"))
cells <-  openSheet("T3")
fillCells(cells, 6 , 2 , mat) 
fillCells(cells, 6 , 10 , mat,names=TRUE)
```


```{r insidefunction, eval=TRUE}
testfun <- function() {
  whichage <- 5
  othervalue <- 8
  fillCells(cells, 10 , 12 , whichage, othervalue )
  
}

testfun()


```


And you can save the resulting Excel file using the saveXlsx function

```{r saving }

saveXlsx()

```

Or if you want to change the name 

```{r saving2}

saveXlsx(wb=NULL,filename = "excelfile2.xlsx")
file.remove("excelfile2.xlsx")

```



