---
title: "xlsxUtils"
author: "epi-gde"
date: "1/24/2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{xlsxUtils}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(epiuf)

```

## xlsxUtility functions

Using xlsx to write simple values to cells may be repetitive and trivial. This set of functions should simplify a lot the process.

The first available function is **openXlsx**. This function load a workbook in memory. The workbook is a java object and having it in environment does help so much... Then this workbook is stored into the functions them self.

It is also possible to assign the workbook to a local variable if needed (to work with more than one workbook at a time).

```{r data}
xls_file <- externalFile("excelfile.xlsx")
openXlsx(xls_file)

```

## the openSheet() function

This function get the content of one sheet from the previously loaded workbook.

```{r usage }

cells <- openSheet("T2")
```

Optionnaly you can load cells from a local workbook.

```{r usage2 }
xls_file <- externalFile("excelfile.xlsx")
wb <- openXlsx(xls_file)
cells <-  openSheet("T2",wb)

```

And finally the function fillCells can be used to input values into a row of cells.

```{r full exemple }
num <-  67
denum <-  90 
cells <- openSheet("T2")
fillCells(cells, 2 , 2 , "Cases","Perc")
fillCells(cells, 3 , 2 , num, num/denum )
cells <- openSheet("T3")
fillCells(cells, 4 , 2 , "total",100)


```

It's also possible to colorize some cells using styles. 
fontColour must be one of the color returned by colours().


```{r colorize }
value <-  5
cells <- openSheet("T2")

# We define two style  and a default style ! 
# In order to be abble to apply defaultStyle in case a specific style was applied before
myStyle <- createXlsxStyle()
italicStyle <- createXlsxStyle(textDecoration = "Italic", fontColour = "green")
alertStyle =  createXlsxStyle(textDecoration = "Bold", fontColour = "red")
boldStyle = createXlsxStyle(textDecoration = "Bold")

fillCells(cells, 17, 1,"Total", style = alertStyle)
fillCells(cells, 17, 2, "Cases", style = italicStyle)
if (value > 10) {
  myStyle <-  alertStyle
} 
fillCells(cells, 17 , 3 , value , style = myStyle)

# if you want erase a previous formatand use wb default : 
myStyle  <- createXlsxStyle()
formatCells(cells, 17, 1, myStyle)

```




```{r matrice}
mat <- data.frame(Id = 1:3 , Vaccs = c("1", "3", "6"))
cells <-  openSheet("T3")
fillCells(cells, 6 , 2 , mat) 
fillCells(cells, 6 , 10 , mat, names=TRUE)
```


```{r insidefunction, eval=TRUE}
testfun <- function() {
  whichage <- 5
  othervalue <- 8
  fillCells(cells, 10 , 12 , whichage, othervalue)
  
}

testfun()


```


And you can save the resulting Excel file using the saveXlsx function

```{r saving }
saveXlsx()

```

Or if you want to change the name 

```{r saving2}

saveXlsx(filename = "excelfile2.xlsx")
# comment the next line if you want to see the result of this vignette
file.remove("excelfile2.xlsx")

```



